---
title: "EU CAYAS NET WP5: EDI Survey recruitment progress"
output: pdf_document
date: "Last update: `r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(patchwork)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
raw_data <- read_csv("data/results-survey632997.csv")

raw_data$startdate<-as.Date(raw_data$startdate, '%Y-%m-%d')

data <- raw_data %>%
  select_all(tolower) %>%
  rename_with(~gsub("0", "_", .x, fixed = TRUE)) %>%
  rename_with(~gsub("[", "_", .x, fixed = TRUE)) %>%
  rename_with(~gsub("]", "", .x, fixed = TRUE)) %>%
  filter(consent_1== "Y")

data$startdate<-as.Date(data$startdate, '%Y-%m-%d')## Keep just day, month, year

data_complete <- data %>%
  filter(lastpage %in% c(8:10)) %>%
  mutate(age_dx = as.numeric(age_dx))
```

* This update covers the period between `r min(raw_data$startdate, na.rm = TRUE)` and `r max(raw_data$startdate, na.rm=TRUE)`.

* Visits to the front page of the survey: `r nrow(raw_data)`.

* Number of participants who consented to participate: `r nrow(data)`.

* Number of participants who completed* the survey: **`r raw_data %>% filter(lastpage %in% c(8:10)) %>% nrow()`** or `r round((raw_data %>% filter(lastpage %in% c(8:10)) %>% nrow())/(data %>% nrow)*100,1)`% of those who consented.

*NB: Completion is determined by having reached the final page of the survey. Front page information stated that data from participant with partial responses will not be used.


## Demographic information of participants who *completed* the survey

## Languages

Participants started the survey in the following languages:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
data_complete %>%
  dplyr::select(startlanguage) %>%
  group_by(startlanguage) %>%
  drop_na() %>%
  count() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         prop = n/sum*100) %>%
  mutate(entry = as.factor(startlanguage),
         entry = fct_reorder(startlanguage, prop)) %>%
  ggplot(aes(x = prop, y = startlanguage, fill = entry)) + 
  geom_col() + 
  geom_text(aes(label = paste0("N = ", n), x = prop+5)) +
  labs(title = " ", 
       x = "Proportion (%)",
       y = " ") + 
  xlim(0, 100) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold'),
        panel.grid.major.x = element_line())
```

## Countries

To date, participants have been recruited from the following countries. NB: The codes are ISO-3 country codes; [source](https://www.iban.com/country-codes). 

OTH = Other; individuals may either wish not to respond, or have a nationality of an unlisted country

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_complete %>%
  dplyr::select(dem_country) %>%
  drop_na() %>%
  group_by(dem_country) %>%
  count() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         prop = n/sum,
         prop = round(prop, 2)) %>%
  dplyr::select(-sum) %>%
  kable(col.names = c("Country code", "N", "Proportion")) %>%
  kable_styling("hover", full_width = FALSE, latex_options = "HOLD_position")
```

## Current age

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3}
data_complete %>%
  ggplot(aes(x = dem_age)) +
  geom_histogram(bins = 30, color = "black", fill="lightblue") +
  #geom_histogram(aes(y=..density..), bins = 30, color = "black", fill="lightblue") +
  #geom_density(fill = "#FF6666", alpha = 0.2) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold')) +
  scale_x_continuous(breaks = seq(0, 80, by=2)) +
  scale_y_continuous(breaks = seq(0, 5, by=1)) +
  labs(x = "Current Age",
       y = "Count")

```


## Age at diagnosis
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=3}
data_complete %>%
  ggplot(aes(x = age_dx)) +
  geom_histogram(bins = 30, color = "black", fill = "lightblue") +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold')) +
  scale_x_continuous(breaks = seq(0, 80, by=2)) +
  scale_y_continuous(breaks = seq(0, 5, by=1)) +
  labs(x = "Current Age",
       y = "Count")
```

## Gender
```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_complete %>%
  dplyr::select(dem_gender) %>%
  mutate(dem_gender = case_when(dem_gender==1 ~ "Female",
                                dem_gender==2 ~ "Male",
                                dem_gender==3 ~ "Non-binary",
                                dem_gender==4 ~ "Other",
                                dem_gender==5 ~ "Prefer not to answer")) %>%
  drop_na() %>%
  group_by(dem_gender) %>%
  count() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         prop = n/sum,
         prop = round(prop, 2)) %>%
  dplyr::select(-sum) %>%
  kable(col.names = c(" ", "N", "Proportion")) %>%
  kable_styling("hover", full_width = FALSE, position = "left",
                latex_options = "HOLD_position")
```

## Treatment status

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_complete %>%
  dplyr::select(tx_now) %>%
  mutate(tx_now = case_when(tx_now==1 ~ "In treatment",
                            tx_now==2 ~ "Not in treatment")) %>%
  drop_na() %>%
  group_by(tx_now) %>%
  count() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         prop = n/sum,
         prop = round(prop, 2)) %>%
  dplyr::select(-sum) %>%
  kable(col.names = c(" ", "N", "Proportion")) %>%
  kable_styling("hover", full_width = FALSE, position = "left",
                latex_options = "HOLD_position")
```

\newpage

# Next steps

* Determine if other info is needed for the updates
* Remove those not fitting the criteria (age)
* Prepare a code for extracting qualitative answers (create a data dictionary)

### Qualitative answers

```{r, results=FALSE}
data_qual <- data_complete %>%
  dplyr::select(id, # To make sure responses aren't double-counted
                startlanguage, # Starting language to know what to translate from
                country_tx_a, tx_finance_qual, 
                s1_2, s1_3, # Secion 1: EDI Statement
                s2_3, s2_2b, # Section 2.1: Bias free environment/care
                s2_5_a, s2_5_b, # Section 2.2: Personal experience 
                s3_2, s3_3) # Section 3: EDI Champions

# write_csv(data_qual, "data/qual_responses.csv")
```


This document was prepared by UK - [reach me with any questions/comments](mailto:dr.urska.kosir@gmail.com)!