---
title: "EU CAYAS NET WP5: EDI Survey recruitment progress"
output: pdf_document
date: "Last update: `r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(patchwork)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
raw_data <- read_csv("data/results-13072023.csv")

raw_data$startdate<-as.Date(raw_data$startdate, '%Y-%m-%d')

data <- raw_data %>%
  select_all(tolower) %>%
  rename_with(~gsub("0", "_", .x, fixed = TRUE)) %>%
  rename_with(~gsub("[", "_", .x, fixed = TRUE)) %>%
  rename_with(~gsub("]", "", .x, fixed = TRUE)) %>%
  filter(consent_1== "Y")

data$startdate<-as.Date(data$startdate, '%Y-%m-%d')## Keep just day, month, year

data_complete <- data %>%
  filter(lastpage %in% c(8:10)) %>%
  mutate(age_dx = as.numeric(age_dx))
```

* This update covers the period between `r min(raw_data$startdate, na.rm = TRUE)` and `r max(raw_data$startdate, na.rm=TRUE)`.

* Visits to the front page of the survey: `r nrow(raw_data)`.

* Number of participants who consented to participate: `r nrow(data)`.

* Number of participants who completed* the survey: **`r raw_data %>% filter(lastpage %in% c(8:10)) %>% nrow()`** or `r round((raw_data %>% filter(lastpage %in% c(8:10)) %>% nrow())/(data %>% nrow)*100,1)`% of those who consented.

*NB*: Completion is determined by having reached the final page of the survey. Front page information stated that data from participant with partial responses will not be used.


## Demographic information of participants who *completed* the survey

## Languages

Participants started the survey in the following languages:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
data_complete %>%
  dplyr::select(startlanguage) %>%
  group_by(startlanguage) %>%
  drop_na() %>%
  count() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         prop = n/sum*100) %>%
  mutate(entry = as.factor(startlanguage),
         entry = fct_reorder(startlanguage, prop)) %>%
  ggplot(aes(x = prop, y = startlanguage, fill = entry)) + 
  geom_col() + 
  geom_text(aes(label = paste0("N = ", n), x = prop+5)) +
  labs(title = " ", 
       x = "Proportion (%)",
       y = " ") + 
  xlim(0, 100) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold'),
        panel.grid.major.x = element_line())
```

## Countries

To date, participants have been recruited from the following countries. NB: The codes are ISO-3 country codes; [source](https://www.iban.com/country-codes). 

OTH = Other; individuals may either wish not to respond, or have a nationality of an unlisted country

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_complete %>%
  dplyr::select(dem_country) %>%
  drop_na() %>%
  group_by(dem_country) %>%
  count() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         prop = n/sum,
         prop = round(prop, 2)) %>%
  dplyr::select(-sum) %>%
  kable(col.names = c("Country code", "N", "Proportion")) %>%
  kable_styling("hover", full_width = FALSE, latex_options = "HOLD_position")
```

## Current age

*NB*: Because the purpose of this document is to inform the recruitment efforts and targets, all cases are included. Further analysis will only be based on the inclusion/exclusion criteria regarding the demographic characteristics such as age range (<39y at diagnosis).

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3}
data_complete %>%
  ggplot(aes(x = dem_age)) +
  geom_histogram(bins = 30, color = "black", fill = "lightblue") +
  #geom_histogram(aes(y=..density..), bins = 30, color = "black", fill="lightblue") +
  #geom_density(fill = "#FF6666", alpha = 0.2) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold')) +
  scale_x_continuous(breaks = seq(0, 80, by=2)) +
  scale_y_continuous(breaks = seq(0, 5, by=1)) +
  labs(x = "Current Age",
       y = "Count")

```

## Age at diagnosis
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=3}
data_complete %>%
  ggplot(aes(x = age_dx)) +
  geom_histogram(bins = 30, color = "black", fill = "lightblue") +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold')) +
  scale_x_continuous(breaks = seq(0, 80, by=2)) +
  scale_y_continuous(breaks = seq(0, 5, by=1)) +
  labs(x = "Current Age",
       y = "Count")
```

## Demographics

```{r echo=FALSE, warning=FALSE}
table_1_dems <- data_complete %>%
  dplyr::select(dem_age, dem_gender, dem_religion, dem_edu, dem_relationship, dem_ses, dem_employment, dem_child, dem_dependent, cancer_dx, tx_now, tx_since, chronic_tx) %>%
  mutate(dem_gender = factor(dem_gender,
                          levels = c(1, 2, 3, 4, 999),
                          labels = c("Female", "Male", "Non-binary", "Other", "Prefer not to answer")),
         dem_religion = factor(dem_religion, 
                               levels = c(0:8, 999),
                               labels = c("None", "Roman Chatholic", "Orthodox", "Protestant", "Jewish", "Hindu", "Muslim", "Buddhist", "Other", "Prefer not to answer")),
         dem_edu = factor(dem_edu,
                          levels = c(1:7),
                          labels = c("Elementary school", "High School", "Higher degree, below university", "Bachelor's", "Master's", "Doctorate/PhD", "Prefer not to answer")),
         dem_relationship = factor(dem_relationship,
                          levels = c(1:4),
                          labels = c("Single", "Married/Common-Law", "Separated/Divorced/Widowed", "Prefer not to answer")),
         dem_ses = factor(dem_ses,
                          levels = c(1:3),
                          labels = c("Low-income", "Middle-income", "High-income")),
         dem_child = factor(dem_child, 
                               levels = c(1, 2, 999),
                               labels = c("Yes", "No", "Prefer not to answer")),
         dem_dependent = factor(dem_dependent, 
                               levels = c(1, 2, 999),
                               labels = c("Yes", "No", "Prefer not to answer")),
         cancer_dx = factor(cancer_dx,
                          levels = c("other", "bladd", "brain", "breas", "cerv", "colre", "leuk", "lung", "mela", "pancr", "prost", "sarco", "testi", "thyro"),
                          labels = c("Other", "Bladder, kidney, liver", "Brain or CNS", "Breast", "Cervical, ovarian, endometrial", "Colon, rectal, gastric", "Hematologic", "Lung", "Melanoma", "Pancreatic", "Prostate", "Sarcoma (all types)", "Testicular", "Thyroid")),
         tx_now = factor(tx_now, 
                               levels = c(1:2),
                               labels = c("Yes", "No")),
         tx_since = factor(tx_since,
                          levels = c(1:3),
                          labels = c("Yes", "No", "Prefer not to answer")),
         chronic_tx = factor(chronic_tx, 
                               levels = c(1:3),
                               labels = c("Yes", "No", "Prefer not to answer")))


table1::label(table_1_dems$dem_gender) <- "Gender"
table1::label(table_1_dems$dem_edu) <- "Education"
table1::label(table_1_dems$dem_religion) <- "Religious orientation"
table1::label(table_1_dems$dem_relationship) <- "Relationship status"
table1::label(table_1_dems$dem_ses) <- "Perceived relative SES"
table1::label(table_1_dems$dem_child) <- "Have children?"
table1::label(table_1_dems$dem_dependent) <- "Care for dependent parents/person?"
table1::label(table_1_dems$cancer_dx) <- "Cancer type"
table1::label(table_1_dems$tx_now) <- "In treatment"
table1::label(table_1_dems$tx_since) <- "Time since treatment"
table1::label(table_1_dems$chronic_tx) <- "Multimorbidity"

table1::table1(~ dem_gender + dem_religion + dem_edu + dem_relationship + dem_ses,
       data = table_1_dems)

table1::table1(~ cancer_dx + tx_now + tx_since + chronic_tx,
       data = table_1_dems)
```

## Pattern of times and missing data

What page are people dropping off at?

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3}
raw_data %>%
  ggplot(aes(x = lastpage)) +
  geom_histogram(bins = 10, color = "black", fill = "lightblue") +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold')) +
  scale_x_continuous(breaks = seq(-1, 10, by=1)) +
  geom_text(aes(y = -1), label = "Last page", x = 8, size = 3) +
  geom_text(aes(y = -1), label = "Consent", x = 1, size = 3) +
  geom_text(aes(y = -1), label = "Med info", x = 3, size = 3) +
  geom_text(aes(y = -1), label = "Sect. 1", x = 4, size = 3) +
  labs(x = "Page where they dropped off",
       y = "How many people (count)")
```


On average participants engaged with the survey for `r round((mean(raw_data$interviewtime))/60,1)` minutes (range `r round((min(raw_data$interviewtime))/60,1)` - `r round((max(raw_data$interviewtime))/60,1)`).

**Those who completed** the survey engaged for an average of **`r round((mean(data_complete$interviewtime))/60,1)`** minutes (range `r round((min(data_complete$interviewtime))/60,1)` - `r round((max(data_complete$interviewtime))/60,1)`).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data_complete %>%
  mutate(interviewtime_min = interviewtime/60) %>%
  filter(interviewtime_min < 60) %>%
  ggplot(aes(x = interviewtime_min)) +
  geom_histogram(bins = 30, color = "black", fill = "lightblue") +
  # geom_histogram(aes(y =..density..), bins = 30, color = "black", fill="lightblue") +
  # geom_density(fill = "#FF6666", alpha = 0.2) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold')) +
  labs(title = "Distribution of time spent filling out the survey",
       x = "Time (min)",
       y = "How many people (count)")
```
\newpage

# Next steps

* export the qualitative data & translate non-English answers

### Qualitative answers

```{r, results=FALSE}
data_qual_1 <- data_complete %>%
  dplyr::select(id, # To make sure responses aren't double-counted
                startlanguage, # Starting language to know what to translate from
                country_tx_a, tx_finance_qual) 

data_qual_2 <- data_complete %>%
  dplyr::select(id, startlanguage,
                s1_2, s1_3) # Section 1: EDI Statement

data_qual_3 <- data_complete %>%
  dplyr::select(id, startlanguage,
                s2_3, s2_2b, # Section 2.1: Bias free environment/care
                s2_5_a, s2_5_b) # Section 2.2: Personal experience 

data_qual_4 <- data_complete %>%
  dplyr::select(id, startlanguage,
                s3_2, s3_3) # Section 3: EDI Champions

# write_csv(data_qual_1, "data/qual_responses_dems.csv")
# write_csv(data_qual_2, "data/qual_responses_sec1.csv")
# write_csv(data_qual_3, "data/qual_responses_sec2.csv")
# write_csv(data_qual_4, "data/qual_responses_sec3.csv")

### ALL:
data_qual <- data_complete %>%
  dplyr::select(id, # To make sure responses aren't double-counted
                startlanguage, # Starting language to know what to translate from
                country_tx_a, tx_finance_qual, 
                s1_2, s1_3, # Secion 1: EDI Statement
                s2_3, s2_2b, # Section 2.1: Bias free environment/care
                s2_5_a, s2_5_b, # Section 2.2: Personal experience 
                s3_2, s3_3) # Section 3: EDI Champions

# write_csv(data_qual, "data/qual_responses.csv")
```


This document was prepared by UK - [reach me with any questions/comments](mailto:dr.urska.kosir@gmail.com)!